An efficient acyclic contact planner for multiped robots

### 2016-01-26 body/equilibrium.tex
typos
finished pass with andrea

% !TEX root =  ../main.tex
\section{Heuristics for relevant contact selection}
\label{sec:heuristics}
\subsection{A heuristic for robust static equilibrium}
The planner is designed so that any generated contact configuration is in static equilibrium.
Equilibrium is of course critical in multipedal locomotion. For this reason we are interested in a robust 
criterion, that ensures that the robot remains balanced in a real-world application, regardless of perception and control uncertainties.

We first give a linear program (LP) that asserts that a contact configuration is in static equilibrium.
From this formulation we derive a new LP that quantifies the robustness of the configuration regarding equilibrium.
In turn, from this value we can either define a heuristic to choose the most robust candidate, or set a threshold on the required robustness. While the presented LP is original, it is based on an analysis of the problem that we proposed in \citep{Prete2016}, where the interested reader can find more details.

\subsubsection{Conditions for static equilibrium.}
We first define the variables of the problem, for a robot $R$ and $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
\item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $\mathbf{m} \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{t}_{i1},\mathbf{t}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{t}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

According to Coulomb's law, the non slipping condition is verified if all the contact forces lie in the friction cone defined by the surface.
As classically done, we linearize the friction cone in a conservative fashion with a pyramid, described by four generating rays of unit length. We choose for instance:
\begin{equation*}
\mathbf{V}_{i} = \mat{\mathbf{n}_{i} + \mu \mathbf{t}_{i1} & \mathbf{n}_{i} -\mu \mathbf{t}_{i1} & \mathbf{n}_{i} + \mu \mathbf{t}_{i2} & \mathbf{n}_{i} - \mu \mathbf{t}_{i2}}^T
\end{equation*}

@@

\begin{equation*}
\forall i  \qquad  \exists \bm{\beta}_i \in \mathbb{R}^{4} : \bm{\beta}_i \ge 0 \wedge \mathbf{f}_{i} = \mathbf{V}_{i} \bm{\beta}_i,
\end{equation*}
where $\bm{\beta}_i$ contains the coefficient of the cone generators.
We can then stack all the constraints to obtain:
\begin{equation}\label{eq:gen}
\exists \bm{\beta} \in \mathbb{R}^{4e} ,  \bm{\beta} \ge 0 \wedge \mathbf{f} = \mathbf{V} \bm{\beta},
\end{equation}
where $\mathbf{V} = \diag{ \{\mathbf{V}_1, \dots, \mathbf{V}_e\} }$, and $\mathbf{f} = (\mathbf{f}_0,...,\mathbf{f}_e)$.

From the Newton-Euler equations, and assuming a null acceleration of the COM, we want the contact forces to compensate the gravity at the center of mass:

@@

If there exists a $\bm{\beta}^*$ satisfying \eqref{eq:gen} and \eqref{eq:new_eul}, it means that the configuration is statically balanced.
The problem can then be formulated as a LP:

@@

We observe that if $b_0$ is positive then \eqref{eq:lin_prog} admits a solution, and $b_0$ is proportional to the minimum distance of the equilibrium contact forces to the boundaries of the friction cones.
If $b_0$ is negative, the configuration is not in static equilibrium, and $b_0$ indicates ``how far'' from equilibrium the configuration is. We thus use $b_0$ as a measure of robustness.

@@

In conclusion, the LP~\eqref{eq:dual} provides a method to assert rapidly static equilibrium, and returns a robustness value.
This robustness criterion can be immediately transformed into a heuristic, and eventually be pondered with others.
Another option to enforce robustness is to set a minimum acceptable value for $b_0$, but as a result this conservative approach is not complete.

>>>>

% !TEX root =  ../main.tex
\section{Heuristics for contact selection}
\label{sec:heuristics}
\subsection{A heuristic for robust static equilibrium}
The planner is designed so that any generated contact configuration is in static equilibrium.
%~ Equilibrium is of course critical in multipedal locomotion. For this reason 
We are interested in a robust 
criterion, that ensures that the robot remains in equilibrium in a real-world application, regardless of perception and control uncertainties.

We first give a linear program (LP) that verifies whether a contact configuration allows for static equilibrium.
From this formulation we derive a new LP that quantifies the robustness of the equilibrium to uncertainties in the contact forces.
In turn, from this value we can either choose the most robust candidate, or set a threshold on the required robustness. While the presented LP is original, it is based on an analysis of the problem that we proposed in \citep{Prete2016}, where the interested reader can find more details.

\subsubsection{Conditions for static equilibrium.}
We first define the variables of the problem, for $e$ contact points, expressed in world coordinates:
\begin{itemize}
\item $\mathbf{c} \in \mathbb{R}^3$ is the robot center of mass (COM);
%~ \item $\mathbf{L}  \in \mathbb{R}^3 $ is the angular momentum at the COM;
\item $\mathbf{m} \in \mathbb{R}$ is the robot mass;
\item $\mathbf{g} = [0,0,-9.81]^T$ is the gravity acceleration;
\item $\mu$ is the friction coefficient;
\item for the i-th contact point $1 \leq i \leq e$:
	\begin{itemize}
	\item $\mathbf{p_i}$ is the contact position;
	\item $\mathbf{f_i}$ is the force applied at $\mathbf{p_i}$;
	\item $\mathbf{n}_i,\mathbf{t}_{i1},\mathbf{t}_{i2}$ form a local Cartesian coordinate system centered at $\mathbf{p_i}$. $\mathbf{n}_i$ is aligned
	with the contact surface normal, and the $\mathbf{t}_i$s are tangent vectors.
	\end{itemize}
\end{itemize}

According to Coulomb's law, the nonslipping condition is verified if all the contact forces lie in the friction cone defined by the surface.
As classically done, we linearize the friction cone in a conservative fashion with a pyramid, described by four generating rays of unit length. We choose for instance:
\begin{equation*}
\mathbf{V}_{i} = \mat{\mathbf{n}_{i} + \mu \mathbf{t}_{i1} & \mathbf{n}_{i} -\mu \mathbf{t}_{i1} & \mathbf{n}_{i} + \mu \mathbf{t}_{i2} & \mathbf{n}_{i} - \mu \mathbf{t}_{i2}}^T
\end{equation*}

@@

\begin{equation*}
\forall i  \qquad  \exists \bm{\beta}_i \in \mathbb{R}^{4} : \bm{\beta}_i \ge 0 \text{ and } \mathbf{f}_{i} = \mathbf{V}_{i} \bm{\beta}_i,
\end{equation*}
where $\bm{\beta}_i$ contains the coefficients of the cone generators.
We can then stack all the constraints to obtain:
\begin{equation}\label{eq:gen}
\exists \bm{\beta} \in \mathbb{R}^{4e} ,  \bm{\beta} \ge 0 \text{ and } \mathbf{f} = \mathbf{V} \bm{\beta},
\end{equation}
where $\mathbf{V} = \diag{ \{\mathbf{V}_1, \dots, \mathbf{V}_e\} }$, and $\mathbf{f} = (\mathbf{f}_0,...,\mathbf{f}_e)$.

From the Newton-Euler equations, to be in static equilibrium the contact forces have to compensate the gravitational forces:

@@

If there exists a $\bm{\beta}^*$ satisfying \eqref{eq:gen} and \eqref{eq:new_eul}, it means that the configuration is in static equilibrium.
The problem can then be formulated as an LP:

@@

We observe that if $b_0$ is positive then \eqref{eq:lin_prog} admits a solution, and $b_0$ is proportional to the minimum distance of the contact forces to the boundaries of the friction cones.
If $b_0$ is negative, the configuration is not in static equilibrium, and $b_0$ indicates ``how far'' from equilibrium the configuration is. We thus use $b_0$ as a measure of robustness.

@@

%~ In conclusion, the LP~\eqref{eq:dual} provides a method to check rapidly static equilibrium, and returns a robustness value.
%~ This robustness criterion can be immediately transformed into a heuristic, and eventually be weighted with others.
%~ Another option to enforce robustness is to set a minimum acceptable value for $b_0$, but as a result this conservative approach discards valid solutions.

### 2016-01-26 body/results.tex
finished pass with andrea

We say that the planning is \gls{interactive} when the computation time for one step is lesser than the
time to execute it. We arbitrarily approximate this time to one second.

@@

\subsection{Description of the scenarios}
In all the scenarios considered, the formulation of the problem is always the same:
a start and goal root configurations are provided as an input of the scenario.
The framework computes the initial contact configuration, and outputs a sequence of contact configurations connecting it to the goal.
In each scenario we detail the parameters chosen: the heuristics, and the constraints on the reachable workspaces (for instance in all the scenarios,
the reachable workspaces of the legs of HRP-2 are always required to intersect with the environment). 
The companion video presents the complete contact sequence obtained in all these scenarios.

@@

\noindent\textbf{Heuristics:} The extended manipulability $h_w$ is chosen for the feet; $h_{EFORT}$ is chosen for the right arm.
Regarding equilibrium, the video demonstrates two sequences computed for two different threshold values of $b_0$: $0$ and $2$ (Figure~\ref{fig:stair_robust}).

@@

Again, the interpolation motion between the contact stances is out of the scope of this paper. However it should be noted that the computed plan in this scenario has been executed successfully on the robot~\citep{Carpentier2016}.

@@

\subsubsection{HRP-2 -- Standing up (Figure~\ref{fig:standing}):}
From a bent configuration, a standing-up motion is computed in a cluttered environment.
The resulting motion involves using a wall as support, and climbing a 25-cm high step.

@@

\noindent\textbf{Observations:} The difficulty of this scenario lies in the strong reduction of the reachable workspace induced 
by the extreme proximity of all obstacles. The planner is able to find a sequence, that consists in many stances.
The proximity of the obstacles invalidate a large number of contact candidates because of collisions. To respect the condition
that only one contact must be broken between each phase, the motion has to be decomposed into a large number of stances (61 in average).
While this scenario is the slower to solve, the planner still computes a solution interactively.
 % \adnote{Maybe define what you mean by 'interactive' and 'real-time' at some point.}

@@

\noindent\textbf{Observations:} In this context, setting up a really important minimum value for $b_0$ is possible due to the high
stability of the HyQ robot, and results in more contact switches, in exchange for safety. The path-planning time is higher than for the previous HRP-2 robots due to the constraint that the 4 reachable workspace of all legs must
collide with the environment at all times. %\adnote{Why do not relax this constraint and require only 3 legs?} 
Again, the computed time remains however \gls{interactive}.

\subsubsection{HyQ -- Obstacle race (Figure~\ref{fig:HyQ_bridge} and~\ref{fig:HyQ_obs}):}
In this long scene, HyQ is first required to cross a 55-cm large hole; then, to cross a narrow ``bridge'', large of only 25-cm.

@@

\noindent\textbf{Heuristics:} $h_w$ for all legs. The robustness threshold $b_0$ is set to $4$.

\noindent\textbf{Observations:} Despite the apparent simplicity of the scene, this scenario is a hard case for a contact planner.
While finding a guide path above the hole is easy for the guide planner, finding a sequence of contacts that allows for equilibrium is not trivial.
Second, the narrow bridge is hard both for the planner and the contact generator: to make sure that equilibrium is preserved along the traversal,
the bridge must be approached with the right angle.
The difficulty is illustrated in Figure~\ref{fig:HyQ_obs}, where several feet rearrangements are required to cross the hole (although the video shows this best).
The planner however succeeds in finding a feasible sequence in the end, again with \gls{interactive} computation times.

@@

\subsubsection{3-fingered hand -- Manipulation of a pen (Figure~\ref{fig:penrot}):}
This scenario is proposed to illustrate the generality of our approach: we consider a manipulation task for a robotic hand and use
our contact planner to compute a guide path for the fingers, considered as effectors (Figure~\ref{fig:penrot}).
Although we do not address the hard issue of accounting for rolling motions, the planner is able to compute the shown sequences in less than 5 seconds.

@@

\subsection{Influence of the parameters} \label{sec:influence}
We discuss the several factors can influence the outcome of our planner: the root scaling factor $s$, the choice of heuristics for
contact generation, the parametrization of the static equilibrium robustness, and lastly, the discretization 
of the guide path.

\subsubsection{Choosing the scaling factor $s$} \label{sec:params}
To find a convenient value $s^*$ of $s$, we proceeded as follows. 
For several values of $s$, we generated $10 000$ configurations. 
We then computed the rates of false positives (ie the resulting configuration satisfies
the reachability condition, but does not belong to \gls{ccontact0}) and false negatives (ie the resulting configuration does not satisfy the 
reachability condition, but belongs to \gls{ccequil0}).

The obtained results for HRP-2 are summarized in Table~\ref{tab:scale}, averaged over all scenes (except for the car egress: in this scenario, 
statistical tests are not really conclusive since we are only interested in a small area of the environment).
As it can be expected, the scaling results in a high increase of the false negative, while the false positive decreases.
For HRP-2 we decided to set $s^*=1.2$, that results in less than 3 \% of false positives, and was also effective
for the car egress scenario.

@@

\subsubsection{On the choice of heuristics} \label{sec:heuristichoices}
In our conference paper, the computed motions were only generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
Regarding locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation on the real robot turned out to be harder \citep{Carpentier2016}.
For this reason, we now prefer to use our manipulability based heuristic for the legs of the robot, but we still
use EFORT for the arm, that results in less contact repositionning.

\subsubsection{parametrization of the robustness equilibrium criterion} \label{sec:parrob}
Robustness is really important when considering practical applications on the robot, to account 
for the various uncertainties that result from environment and state estimator. However,
maximizing the robustness criterion is not necessarily optimal, because the resulting contacts will stay close to the center
of mass, thus not favoring the motion. In our experiments, we choose not to maximize the robustness,
but to empirically set a robustness threshold value under which a configuration is not considered to be in
static equilibrium (If no candidate reaches the threshold, instead of failing, the algorithm can eventually return the ``more robust'' configuration
found). Currently for HRP-2, a threshold value of 2 subjectively gives the best results in the considered scenarios, while 10 seems to be a good choice for HyQ. Both values do not
significantly slow down the planning times.

\subsubsection{Discretization of the guide path} \label{sec:disc}
Currently, to discretize the path, the user defines a fixed step size. The step size
has an influence on the output of the planner: if too large steps are taken,
the planner may fail since we impose the constraint that only one contact change might occur
between two consecutive stances. For instance, in most scenarios the torso of HRP-2 translates by about 15 cm between two postures, but about only 3 cm
for the car egress scenario for instance, where we had to lower the step size because of the many contact breaks, induced by the extremely constrained geometry.
For future work, we would like to automatically adapt the size of the discretization step to the complexity of the planning, for instance
by reducing the step size when in the areas where failure occur.

@@

\subsubsection{Computation times}
Table~\ref{tab:requestime} summarizes the performance measurement obtained, in terms of computation times.

Considering the repartition of the computation time, for HRP-2, most of the time is spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the car egress scenario,
the kinematic constraints are really strong. Therefore a large number of projections fail because of joint limits violation.

For HyQ, there is a more regular repartition of the performances for the rubble scenario. For the obstacle scenario,
we observe a large increase of the time spent performing collision checking and inverse kinematics. This is explained
by the complexity of crossing the bridge: the robot has to stand high and create contacts close to each other because
the bridge is narrow. In many cases this results in collisions, thus a large number of contacts have to be evaluated.

In all scenarios, one can observe that the average computation time for one single step is largely below the second,
thus allowing to consider \gls{interactive} applications.

Table~\ref{tab:requestpercent} summarizes the success rates obtained for each scenario.

\subsubsection{Success rates}
We observe that, as expected, our planner does not succeed systematically, because of the approximations made in the implementation
of our theoretical formulation. The extreme situation of the car egress scenario provides the less satisfying results in terms of kinematic failure. This is not
surprising considering the narrowness of the scene.
From a pragmatic point of view, regarding the computation times, we claim that our approach provides a satisfying compromise between completeness and efficiency.
Indeed, the advantage of the framework is that when the contact generation fails, it does so rapidly, which allows us to rapidly re-plan a new contact sequence with a reasonable chance of success.
The most efficient (and immediate) approach is probably to launch in parallel several instances of the planner for a given problem (our current implementation is single threaded) and to apply any successful result.

\begin{table*}
\centering
\begin{tabular}{ c | c | c | c | c }
 Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{5.42 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours} + \citeauthor{Carpentier2016}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & $ <$ 2s \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{4.08 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & X \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{10.08 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
 \hline
   \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{3.5 hours}  \\
							 & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 21 s} & X \\
 \end{tabular}
\caption{TODO VERIFIE STATS A LA MAISON.}
\label{tab:compprev}
\quad
\end{table*}
\subsection{Comparison with previous work} \label{sec:compa}
Comparing our method with others is not trivial. Indeed, we focus on the contact planning aspect of the problem, and do not address
the interpolation in this paper, nor provide a strong guarantee that our sequence can be interpolated.
Papers in the state of the art, when they indicate the computation times, often only provide an overall computation
time, and we can not afford the important development resources required to reimplement them.

Furthermore, the interpolation itself is a problem that is far from being solved. For instance in \cite{DBLP:conf/iser/EscandeKMG08}, the end-effector trajectories are computed automatically with splines, but are manually corrected if they enter in collision. In \citep{Mordatch:2012:DCB:2185520.2185539}, internal joint collisions can occur in the resulting motion.
Recent methods suffer from similar issues \citep{Carpentier2016}.

Regardless, we indicate in Table~\ref{tab:compprev} what is actually computed by the method and in how much time.
We compare motions obtained in previous works with HRP-2 (or a humanoid avatar) with ours: the two scenarios we selected consist in:  climbing one step, high of 20, 30 or 40 centimeters; 
moving out of a table, that we consider of a complexity similar to the car egress scenario (or inferior since the feet location remain on the ground in the former case).
For scenarios that apply, we also indicate the interpolation time that we obtain by feeding the plan to our interpolation method presented in \cite{Carpentier2016}.

>>>>

%~ We say that the planning is \gls{interactive} when the computation time for one step is lesser than the
%~ time to execute it. We arbitrarily approximate this time to one second.

@@

\subsection{Description of the scenarios}
In all the scenarios considered, the formulation of the problem is always the same:
a start and goal root configurations are provided as an input of the scenario.
The framework computes the initial contact configuration, and outputs a sequence of contact configurations connecting it to the goal.
In each scenario we detail the parameters chosen: the heuristics, and the constraints on the reachable workspaces (for instance in all the scenarios,
the reachable workspaces of the legs of HRP-2 are always required to intersect with the environment). 
%~ The companion video presents the complete contact sequence obtained in all these scenarios.

@@

\noindent\textbf{Heuristics:} The manipulability $h_w$ is chosen for the feet; $h_{EFORT}$ is chosen for the right arm.
Regarding equilibrium, the video demonstrates two sequences computed for two different threshold values of $b_0$: $0$ and $2$ (Figure~\ref{fig:stair_robust}).

@@

Again, the interpolation motion between the contact steps is out of the scope of this paper. However it should be noted that the computed plan in this scenario has been executed successfully on the robot~\citep{Carpentier2016}.

@@

\subsubsection{HRP-2 -- Standing up (Figure~\ref{fig:standing}):}
From a bent configuration, a standing-up motion is computed in a constraining environment.
The resulting motion involves using a wall as support, and climbing a 25-cm high step.

@@

\noindent\textbf{Observations:} The difficulty of this scenario lies in the strong reduction of the reachable workspace induced 
by the extreme proximity of all obstacles. The planner is able to find a sequence, that consists in many steps.
The proximity of the obstacles invalidate a large number of contact candidates because of collisions. To avoid breaking more than one contact between each step, the motion has to be decomposed into a large number of steps (61 in average).
While this scenario is the slowest to solve, the planner still computes a solution \glslink{interactive}{interactively}.
 % \adnote{Maybe define what you mean by 'interactive' and 'real-time' at some point.}

@@

\noindent\textbf{Observations:} In this context, setting up a really important minimum value for $b_0$ is possible due to the high
stability of the HyQ robot, and results in more contact switches, in exchange for safety. The path-planning takes a few seconds due to the constraint that the 4 reachable workspaces of all legs must
collide with the environment at all times. %\adnote{Why do not relax this constraint and require only 3 legs?} 
Again, the computation times remain however \gls{interactive}.

\subsubsection{HyQ -- Obstacle race (Figure~\ref{fig:HyQ_bridge} and~\ref{fig:HyQ_obs}):}
In this long scene, HyQ is first required to cross a 55-cm large hole; then, to cross a narrow ``bridge'', only 25-cm large.

@@

\noindent\textbf{Heuristics:} $h_w$ for all legs. The robustness threshold $b_0$ is set to $10$.

\noindent\textbf{Observations:} Despite the apparent simplicity of the scene, this scenario is a hard case for a contact planner.
While finding a guide path above the hole is easy for the guide planner, finding a sequence of contacts that allows for equilibrium is not trivial.
Second, the narrow bridge is hard both for the planner and the contact generator: to make sure that equilibrium is preserved along the traversal,
the bridge must be approached with the appropriate angle.
The difficulty is illustrated in Figure~\ref{fig:HyQ_obs}, where several feet rearrangements are required to cross the hole (although the video shows this best).
The planner however succeeds in finding a feasible sequence in the end, again with \gls{interactive} computation times.

@@

\subsubsection{3-fingered hand -- Manipulation of a pen (Figure~\ref{fig:penrot}):}
This scenario is proposed to illustrate the generality of our approach: we consider a manipulation task for a robotic hand and use
our contact planner to compute a contact sequence for the fingers, considered as effectors (Figure~\ref{fig:penrot}).
Although we do not address the hard issue of accounting for rolling motions, the planner is able to compute the shown sequences in less than 5 seconds.

@@

\subsection{Influence of the parameters} \label{sec:influence}
This subsection discusses the several factors that influence the outcome of our planner: the root scaling factor $s$, the choice of heuristics for
contact generation, the desired value of the robustness parameter $b_0$, and lastly, the discretization 
of the guide path.

\subsubsection{Choosing the scaling factor $s$} \label{sec:params}
To find a convenient value $s^*$ of $s$, we proceeded as follows. 
For several values of $s$, we generated $10 000$ configurations. 
We then computed the rates of false positives (i.e. the configuration satisfies
the reachability condition, but does not belong to \gls{ccequil0}) and false negatives (i.e. the configuration does not satisfy the 
reachability condition, but belongs to \gls{ccequil0}).

The obtained results for HRP-2 are summarized in Table~\ref{tab:scale}, averaged over all scenes (except for the car egress: in this scenario, 
statistical tests are not really conclusive since we are only interested in a small area of the environment).
As it can be expected, the scaling results in a high increase of the false negatives, while the false positives decrease.
For HRP-2 we decided to set $s^*=1.2$, that results in less than 3 \% of false positives, and was also effective
for the car egress scenario.

@@

\subsubsection{On the choice of heuristics} \label{sec:heuristichoices}
In our conference paper, the computed motions were only generated using the EFORT heuristic.
EFORT is designed for tasks requiring to exert important forces (such as pushing / pulling / climbing). 
Regarding locomotion tasks, such as the stair scenario, one issue with EFORT is that it tends to generate
configurations close to singularities (and joint limits). While this does not significantly impact
the generation of the plan, the resulting interpolation on the real robot turned out to be harder \citep{Carpentier2016}.
For this reason, we now prefer to use our manipulability-based heuristic for the legs of the robot, but we still
use EFORT for the arms, that results in less contact repositioning.

\subsubsection{On the robustness equilibrium criterion} \label{sec:parrob}
Robustness is really important when considering practical applications on the robot, to account 
for the various uncertainties that result from environment and state estimation. However,
maximizing the robustness criterion is not necessarily optimal, because the resulting configurations may be too conservative, thus not favoring the motion. In our experiments, we choose not to maximize the robustness,
but to empirically set a robustness threshold value under which a configuration is not considered to be in
static equilibrium (If no candidate reaches the threshold, instead of failing, the algorithm can eventually return the ``more robust'' configuration
found). Currently for HRP-2, a threshold value of 2 subjectively gives the best results in the considered scenarios, while 10 seems to be a good choice for HyQ. Both values do not
significantly slow down the planning times.

\subsubsection{Discretization of the guide path} \label{sec:disc}
Currently, to discretize the path, the user defines a fixed step size. The step size
has an influence on the output of the planner: if too large steps are taken,
the planner may fail since we impose the constraint that only one contact change might occur
between two consecutive steps. For instance, in most scenarios the torso of HRP-2 moves about 15 cm between two postures, but only 3 cm
for the car egress scenario, where we had to lower the step size because of the many contact breaks induced by the extremely constraining geometry.
For future work, we would like to automatically adapt the size of the discretization step to the complexity of the planning, for instance
by reducing the step size when failure occurs.

@@

\subsubsection{Computation times}
Table~\ref{tab:requestime} summarizes the performance measurements obtained, in terms of computation times.

Considering the repartition of the computation time, for HRP-2, most of the time is spent performing inverse kinematics.
This is not surprising considering the number of calls to the methods: IK projection is used intensively to maintain contact continuity between two postures; 
it is also applied every time a new candidate needs to be evaluated. In particular for the car egress scenario,
the kinematic constraints are really strong. Therefore a large number of projections fail because of joint-limit violations.

For HyQ, there is a more uniform repartition of the times for the rubble scenario. For the obstacle scenario,
we observe a large increase of the time spent performing collision checking and inverse kinematics. This is explained
by the complexity of crossing the bridge: the robot has to stand high and create contacts close to each other because
the bridge is narrow. In many cases this results in collisions, thus a large number of contacts have to be evaluated.

In all scenarios, one can observe that the average computation time for one single step is largely below one second,
thus allowing to consider \gls{interactive} applications.

\subsubsection{Success rates}
Table~\ref{tab:requestpercent} summarizes the success rates obtained for each scenario.
We observe that, as expected, our planner does not succeed systematically, because of the approximations made in our formulation. The extreme situation of the car egress scenario provides the less satisfying results in terms of kinematic failure. This is not
surprising considering the narrowness of the scene.
From a pragmatic point of view, regarding the computation times, we claim that our approach provides a satisfying compromise between completeness and efficiency.
Indeed, the advantage of the framework is that when the contact generation fails, it does so rapidly, which allows us to rapidly re-plan a new contact sequence with a reasonable chance of success.
The most efficient (and immediate) approach is probably to launch in parallel several instances of the planner for a given problem (our current implementation is single threaded) and to apply any successful result.

\begin{table*}
\centering
\begin{tabular}{ c | c | c | c | c }
 Scenario & Method  & Guide Path & Contact sequence & Interpolation \\
 \hline
   \multirow{3}{*}{Stair 20 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{5.42 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours} + \citeauthor{Carpentier2016}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & $ <$ 2s \\
 \hline
   \multirow{3}{*}{Stair 30 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{4.08 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours}  & $\mathbf{<}$ \textbf{2 ms} & $\mathbf{<}$ \textbf{14 ms} & X \\
 \hline
   \multirow{3}{*}{Stair 40 cm} & \citeauthor{Hauser06usingmotion} &  \multicolumn{3}{c}{10.08 min}  \\
							 & \citeauthor{Mordatch:2012:DCB:2185520.2185539} &  \multicolumn{3}{c}{2 to 10 min} \\
							 & \textbf{Ours}  & \textbf{5 s} & \textbf{3 s} & X \\
 \hline
   \multirow{2}{*}{Table (car) egress} & \citeauthor{Bouyarmane2009, DBLP:conf/iser/EscandeKMG08} & 10 min & \multicolumn{2}{c}{3.5 hours}  \\
							 & \textbf{Ours}  & \textbf{529 ms} & \textbf{$<$ 21 s} & X \\
 \end{tabular}
\caption{TODO VERIFIE STATS A LA MAISON.}
\label{tab:compprev}
\quad
\end{table*}
\subsection{Comparison with previous work} \label{sec:compa}
Comparing our method with others is not trivial. Indeed, we focus on the contact planning aspect of the problem, and do not address
the interpolation in this paper, nor provide a strong guarantee that our sequences can be interpolated.
Most papers in the literature only provide the overall computation
time, without specifying how much is spent in each phase. Due to the complexity
of the existing approaches,  we cannot afford the development resources required to reimplement them.

Furthermore, the interpolation itself is not completely solved. For instance in \cite{DBLP:conf/iser/EscandeKMG08}, the end-effector trajectories are computed automatically with splines, but are manually corrected if they enter in collision. In \citep{Mordatch:2012:DCB:2185520.2185539}, internal joint collisions can occur in the resulting motion.
Recent methods suffer from similar issues \citep{Carpentier2016}.

Nonetheless, we indicate in Table~\ref{tab:compprev} what is actually computed by each method and in how much time.
We compare motions obtained in previous works with HRP-2 (or a humanoid avatar) with ours.
The two scenarios we selected consist in:  climbing one step, of height 20, 30 or 40 centimeters; 
getting out of a table, that we consider of a complexity similar to the car egress scenario (or inferior since the feet locations remain on the ground in the former case).
For scenarios that apply, we also indicate the interpolation time that we obtained by feeding the plan to our interpolation method presented in \cite{Carpentier2016}.

### 2016-01-26 main.tex
nico + glossary
complete, need abstract and section 4
typos
finished pass with andrea

\setcounter{secnumdepth}{3}
\usepackage{moreverb,url}
\usepackage{makeidx}         % allows index generation
\usepackage[percent]{overpic}
\usepackage{array} 
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{color}
\usepackage{bm}

@@

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}

@@

@@

\author{Steve Tonneau\affilnum{1}, Andrea Del Prete\affilnum{1}, Julien Pettr\'e\affilnum{2}, Chonhyon Park\affilnum{3}, Dinesh Manocha\affilnum{3}, Nicolas Mansard\affilnum{1}}
 %~ Franck Multon\affilnum{3},
\affiliation{\affilnum{1}LAAS-CNRS, Toulouse, France\\
\affilnum{2}Inria, Rennes, France\\
\affilnum{3}UNC, Chapel Hill, USA  } % \\
%~ \affilnum{3}Universit\'e Rennes 2, France}

@@

\section{Introduction}
\input{body/intro}
\input{body/overview}
\input{body/rbprm}
\input{body/contact}
\input{body/equilibrium}
\input{body/efort}
\input{body/results}
\input{body/conclusion}
\section*{Acknowledgements}
This research is supported by Euroc (project under FP7 Grant Agreement  608849),  Entracte (ANR  grant  agreement  13-CORD-002-01),
the ARO Contract W911NF-14-1-0437,
the NSF award 1305286, and a grant from Boeing.
\theendnotes
\bibliographystyle{SageH}
\bibliography{biblio}
\appendix
\input{body/appendix_roms}
\input{body/appendix_algo}
\input{body/appendix_hpp}
\end{document}

>>>>

\setcounter{secnumdepth}{3}
\usepackage{moreverb,url}
\usepackage{makeidx}         % allows index generation
\usepackage[percent]{overpic}
\usepackage{array} 
\usepackage{algorithm}
\usepackage[noend]{algpseudocode}
\usepackage{listings}
\usepackage{color}
\usepackage{bm}
\usepackage{booktabs}
\usepackage{multirow}

@@

\usepackage[colorlinks,bookmarksopen,bookmarksnumbered,citecolor=red,urlcolor=red]{hyperref}
\usepackage{glossaries}

@@

\makeglossaries
\input{body/glossary}

@@

\author{Steve Tonneau\affilnum{1}, Andrea Del Prete\affilnum{1}, Julien Pettr\'e\affilnum{2}, Chonhyon Park\affilnum{3}, Dinesh Manocha\affilnum{3}, Nicolas Mansard\affilnum{1}}
 %~ Franck Multon\affilnum{3},
\affiliation{\affilnum{1}LAAS-CNRS / Universit\'e Paul Sabatier, Toulouse, France\\
\affilnum{2}Inria, Rennes, France\\
\affilnum{3}UNC, Chapel Hill, USA  } % \\
%~ \affilnum{3}Universit\'e Rennes 2, France}

@@

\section{Introduction}
\input{body/intro}
\input{body/overview}
\input{body/rbprm}
\input{body/contact}
\input{body/equilibrium}
\input{body/efort}
\input{body/results}
\input{body/conclusion}
\section*{Acknowledgements}
This research is supported by Euroc (project under FP7 Grant Agreement  608849),  Entracte (ANR  grant  agreement  13-CORD-002-01),
the ARO Contract W911NF-14-1-0437 and the NSF award 1305286. %, and a grant from Boeing.
\theendnotes

\glsaddall
\printglossaries
\bibliographystyle{SageH}
\bibliography{biblio}
\appendix
\input{body/appendix_roms}
\input{body/appendix_algo}
\input{body/appendix_hpp}
\end{document}
